Other metrics

### 2021-02-25 analysis/rev_prototyping/additional_metrics_results.Rmd
figuregit add --allgit add --all

all_di  %>% 
  group_by(dat) %>%
  filter(nparts > 20) %>%
  summarize(
            prop_shannon_high_raw = mean(shannon_percentile > 95),
            prop_shannon_low_raw = mean(shannon_percentile < 5),
            nsites_included = dplyr::n())
```

@@

>>>>

all_di  %>% 
  group_by(dat) %>%
  filter(nparts > 20) %>%
  summarize(
            prop_shannon_high_raw = mean(shannon_percentile_excl > 95),
            prop_shannon_low_raw = mean(shannon_percentile < 5),
            nsites_included = dplyr::n())
```

@@

## Epic synthesizing figure

```{r, fig.dim = c(5, 10)}

all_di_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"))

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

jk_di <- jk_di_full %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, nsingletons_percentile, nsingletons_percentile_excl, nsingletons, nsingeltons_median, nsingletons_95, mean_po_comparison_percentile, mean_po_comparison_sims, mean_po_comparison_95, skew_95_ratio_1t, simpson_95_ratio_1t, shannon_95_ratio_1t, mean_po_comparison_95_ratio_1t, nparts)

jk_di_means <- jk_di %>%
  group_by(dat, site_source) %>%
  summarise_at(vars(-group_cols(), -site), mean) %>%
  mutate(resampling = "jacknife") %>%
  mutate(site = as.integer(site_source)) %>%
  select(-site_source) %>%
  filter(n0 != s0,
         s0 != 1,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20)

all_di_with_resamples <- all_di_with_resamples %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
       #  !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  mutate(resampling = ifelse(singletons, "rarefaction", "none")) %>%
  bind_rows(jk_di_means)

prop_results <- all_di_with_resamples %>%
 # filter(!singletons) %>%
  group_by(Dataset, resampling) %>%
  summarize(
    skew_high = mean(skew_percentile_excl > 97.5),
    skew_low = mean(skew_percentile < 2.5),
    even_high = mean(simpson_percentile_excl > 97.5),
    even_low = mean(simpson_percentile < 2.5),
    shannon_high = mean(shannon_percentile > 97.5),
    shannon_low = mean(shannon_percentile < 2.5),
    nsingletons_high = mean(nsingletons_percentile_excl > 97.5),
    nsingletons_low = mean(nsingletons_percentile < 2.5),
    po_high = mean(mean_po_comparison_percentile > 97.5),
    po_low = mean(mean_po_comparison_percentile < 2.5)
  ) %>%
  ungroup() %>%
  tidyr::pivot_longer((-c("Dataset","resampling")), names_to = "metric", values_to = "value") %>%
  mutate(ometric = ordered(metric, levels = c("skew_high", "skew_low", "even_low", "even_high", "shannon_low", "shannon_high", "nsingletons_high", "nsingletons_low", "po_high", "po_low")),
         ometric2 = ordered(metric, levels = c("skew_high",  "even_low",  "shannon_low",  "nsingletons_high", "po_high", "skew_low","even_high","shannon_high", "nsingletons_low", "po_low"))) %>%
  mutate(weird_dir = as.numeric(ometric2) > 5)

# 
# ggplot(prop_results, aes(Dataset, value, color = resampling)) +
#   geom_point() +
#   geom_point(data = filter(prop_results, resampling == "none"), color = "black", size = 2) +
#   geom_hline(yintercept = .025) +
#   facet_wrap(vars(ometric), ncol = 2, scales = "free_y") +
#   theme(legend.position = "top")

ggplot(prop_results, aes(ometric2, value, color = resampling)) +
  geom_point() +
  geom_point(data = filter(prop_results, resampling == "none"), color = "black", size = 2) +
  geom_hline(yintercept = .025) +
  facet_grid(rows = vars(Dataset), cols = vars(weird_dir), scales = "free_x", switch = "y") +
  theme(legend.position = "top") +
  theme(axis.text.x = element_text(angle = 75, hjust= 1))
```

This is an EPIC figure.

Specs:

- The row panels are different datasets. 
- The left column of panels is the "usual" direction of the effect - the way most datasets behave. The right column is the "reverse" direction - the way Gentry behaves.
- The x axis is different metrics.
- The y axis is the proportion of sites with an extreme value (> 97.5 or < 2.5 percentile, depending on if the metric is "high" or "low" respectively) for that metric. 
- The horizontal line is 2.5. Because this is now making a 2tailed comparison, the 2.5 cutoff is the proportion of percentile values we would expect to be in this range at random.
- The BLACK points are for raw SADs. The BLUE dots are for SADs corrected for nondetection of rare species. The RED dots are for SADs resampled via jacknife resampling.
- Note that a lot of FIA stuff hasn't finished running, so don't take those panels seriously.

Interpretation:

- All datasets (except maybe FIA) have extreme values more often than usual for almost all metrics.
- Adjusting for rare species always increases this effect if it has any impact at all.
- Jacknife resampling decreases it, but only to no significance for BBS and MCDB # of singletons.
- While Gentry often has extreme values in the same direction as the other datasets, it also has them very strongly in the *opposite* direction.
